---
title: "final_model"
author: "Tam Ta"
date: "5/3/2022"
output: html_document
---
#### To do list
-work on brochures and presentations
-figure out how to make the graphs
-run the models

#### Ideas
-plot models over data - x=use/nonuse (1/0), y=elevation, plot prediction line on top
-boxplots/bar graph to compare mule deer vs elk elevation, slope, canopy cover


#### Setup
```{r load library, include=FALSE}
library(here)
library(tidyverse)
library(dplyr)
library(raster) #for scaling
library(AICcmodavg) #for AIC
library(ggplot2)
library(readxl)
library(car) #for vif
library(MuMIn) #for dredge
library(tibble) #for add_column
library(jtools)
```

```{r load data, include=FALSE}
kill<-read_excel(here("data/killdata.xlsx"))%>%
  distinct(kill_num,.keep_all = TRUE)%>%
  mutate(kill=1,
         canopy_cover=`Canopy cover`)%>%
  dplyr::select(kill,species,year,jday,long,lat,elevation,slope,
                canopy_cover)
random<-read_excel(here("data/randomdata.xlsx"))%>%
  mutate(kill=0,
         slope=Slope,
         canopy_cover=`Canopy cover`)%>%
  dplyr::select(kill,long,lat,elevation,slope,canopy_cover)
```

#### Prepare datasets for model

```{r join kills and random points}
premod<-bind_rows(random,kill)
```

Scale any numerical columns, except lat/long and use/nonuse. This would involve elevation, slope, and canopy cover. Scaling would help standardize the numerical covariates --> allows us to compare coefficients to each other.

```{r scale}
premod.scale <- scale(premod[4:6])

#add back any columns that we did not scale
mod <- data.frame("kill" = premod$kill, "species" = premod$species,
                  premod.scale)
```

```{r filter data, include=FALSE}
elk.mod<-mod%>%
  filter(species=="elk" | kill == 0)
mule.mod<-mod%>%
  filter(species=="mule deer" | kill == 0)
check<-elk.mod%>%
  filter(kill==0)
```

#### Models & AIC

Null model to compare to logistic regressions. If AIC of logistics are above or <2 below null model, then the relationships are not significant. 
```{r null model}
null <- glm(kill ~ 1, family = binomial, data=mod)
summary(null)
```

```{r elk model}
elk<-glm(kill~elevation+canopy_cover+slope, data=elk.mod,
          family=binomial,na.action = "na.fail")
elk1<-glm(kill~elevation+canopy_cover+slope + I(slope^2), data=elk.mod,
          family=binomial,na.action = "na.fail")
elk.elev<-glm(kill~elevation,data=elk.mod,
          family=binomial,na.action = "na.fail")
elk.slope<-glm(kill~slope,data=elk.mod,
          family=binomial,na.action = "na.fail")
elk.cano<-glm(kill~canopy_cover,data=elk.mod,
          family=binomial,na.action = "na.fail")
summary(elk1)
vif(elk)
dredge(elk)
confint(elk)
```

```{r mule deer model}
mule<-glm(kill~elevation+canopy_cover+slope, data=mule.mod,
          family=binomial,na.action = "na.fail")
mule1<-glm(kill~elevation+canopy_cover+slope+I(slope^2), data=mule.mod,
          family=binomial,na.action = "na.fail")
mule.elev<-glm(kill~elevation, data=mule.mod,
          family=binomial,na.action = "na.fail")
mule.slope<-glm(kill~slope, data=mule.mod,
          family=binomial,na.action = "na.fail")
mule.cano<-glm(kill~canopy_cover, data=mule.mod,
          family=binomial,na.action = "na.fail")
summary(mule1)
vif(mule)
dredge(mule)
confint(mule)
```

Lower AIC = better. Choose model with the lowest AIC. Make sure it is >2 below the null.

```{r aic table}
#the models you're comparing
model.set <- list(m.null, m.elk, m.mule)
model.names <- c("null", "elk", "mule deer")

#makes the table
aic = aictab(model.set, modnames = model.names)
aic
```

#### Graphs

```{r prediction tables}
#elevation
mule.plot <- with(mule.mod,
  data.frame(elevation=seq(min(mule.mod$elevation), 
                           max(mule.mod$elevation),len=500)))
mule.plot <- add_column(mule.plot, 
                        fit = predict(mule.elev, newdata=mule.plot,
                                      type = 'response'))
elk.plot <- with(elk.mod,
  data.frame(elevation=seq(min(elk.mod$elevation), 
                           max(elk.mod$elevation),len=500)))
elk.plot <- add_column(elk.plot, 
                        fit = predict(elk.elev, newdata = elk.plot,
                                      type = 'response'))
#slope
mule.plot1 <- with(mule.mod,
  data.frame(slope=seq(min(mule.mod$slope), 
                           max(mule.mod$slope),len=500)))
mule.plot1 <- add_column(mule.plot1, 
                        fit=predict(mule.slope, newdata=mule.plot1,
                                      type = 'response'))
elk.plot1 <- with(elk.mod,
  data.frame(slope=seq(min(elk.mod$slope), 
                           max(elk.mod$slope),len=500)))
elk.plot1 <- add_column(elk.plot1, 
                        fit = predict(elk.slope, newdata=elk.plot1,
                                      type = 'response'))
#canopy cover
mule.plot2 <- with(mule.mod,
  data.frame(canopy_cover=seq(min(mule.mod$canopy_cover), 
                           max(mule.mod$canopy_cover),len=500)))
mule.plot2 <- add_column(mule.plot2, 
                        fit=predict(mule.cano, newdata=mule.plot2,
                                      type = 'response'))
elk.plot2 <- with(elk.mod,
  data.frame(canopy_cover=seq(min(elk.mod$canopy_cover), 
                           max(elk.mod$canopy_cover),len=500)))
elk.plot2 <- add_column(elk.plot2, 
                        fit = predict(elk.cano, newdata=elk.plot2,
                                      type = 'response'))
```

```{r confidence interval}
#elk elevation
ilink <- family(elk.elev)$linkinv
elk.con <- bind_cols(elk.plot, setNames(as_tibble(predict(elk.elev, elk.plot, se.fit = TRUE)[1:2]),c('fit_link','se_link')))
elk.con <- mutate(elk.con,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)),
                species="Elk")
#elk slope
ilink <- family(elk.slope)$linkinv
elk.con1 <- bind_cols(elk.plot1, setNames(as_tibble(
  predict(elk.slope, elk.plot1, se.fit = TRUE)[1:2]),
  c('fit_link','se_link')))
elk.con1 <- mutate(elk.con1,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)),
                species="Elk")
#elk canopy cover
ilink <- family(elk.cano)$linkinv
elk.con2 <- bind_cols(elk.plot2, setNames(as_tibble(
  predict(elk.cano, elk.plot2, se.fit = TRUE)[1:2]),
  c('fit_link','se_link')))
elk.con2 <- mutate(elk.con2,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)),
                species="Elk")
#mule deer elevation
ilink <- family(mule.elev)$linkinv
mule.con <- bind_cols(mule.plot,setNames(as_tibble(predict(
  mule.elev, mule.plot,se.fit = TRUE)[1:2]), 
  c('fit_link','se_link')))
mule.con <- mutate(mule.con,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)),
                species="Mule deer")
#mule deer slope
ilink <- family(mule.slope)$linkinv
mule.con1 <- bind_cols(mule.plot1, setNames(as_tibble(
  predict(mule.slope, mule.plot1, se.fit = TRUE)[1:2]),
  c('fit_link','se_link')))
mule.con1 <- mutate(mule.con1,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)),
                species="Mule deer")
#mule deer canopy cover
ilink <- family(mule.cano)$linkinv
mule.con2 <- bind_cols(mule.plot2, setNames(as_tibble(
  predict(mule.cano, mule.plot2, se.fit = TRUE)[1:2]),
  c('fit_link','se_link')))
mule.con2 <- mutate(mule.con2,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)),
                species="Mule deer")
```

```{r bind tables}
elev.plot<-rbind(mule.con,elk.con)
slope.plot<-rbind(mule.con1,elk.con1)
cano.plot<-rbind(mule.con2,elk.con2)
```

```{r graph prediction tables}
#elevation
ggplot(elev.plot)+
  geom_ribbon(aes(x=elevation, ymin = right_lwr,ymax = right_upr,
                  fill=species),alpha = 0.2)+
  scale_fill_manual(values = c("#39638b","#26a185"),
                    labels=c("Elk","Mule deer"), name="Species")+
  geom_smooth(aes(x=elevation, y=fit, col=species))+
  scale_color_manual(values = c("#39638b","#26a185"),
                     labels=c("Elk","Mule deer"), name="Species")+
  labs(x="Elevation (m)", y="Probability of kill")+
  theme(legend.key=element_rect(fill="white"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))
ggsave(here("figures/elevation.conf.png"),dpi=300)

#slope
ggplot(slope.plot)+
  geom_ribbon(aes(x=slope, ymin = right_lwr,ymax = right_upr,
                  fill=species),alpha = 0.2)+
  scale_fill_manual(values = c("#39638b","#26a185"),
                    labels=c("Elk","Mule deer"), name="Species")+
  geom_smooth(aes(x=slope, y=fit, col=species))+
  scale_color_manual(values = c("#39638b","#26a185"),
                     labels=c("Elk","Mule deer"), name="Species")+
  labs(x="Slope", y="Probability of kill")+
  theme(legend.key=element_rect(fill="white"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))
ggsave(here("figures/slope.conf.png"),dpi=300)

#canopy cover
ggplot(cano.plot)+
  geom_ribbon(aes(x=canopy_cover, ymin=right_lwr, ymax=right_upr,
                  fill=species),alpha = 0.2)+
  scale_fill_manual(values = c("#436658","#b4673b"),
                    labels=c("Elk","Mule deer"), name="Species")+
  geom_smooth(aes(x=canopy_cover, y=fit, col=species))+
  scale_color_manual(values = c("#436658","#b4673b"),
                     labels=c("Elk","Mule deer"), name="Species")+
  labs(x="Canopy cover (%)", y="Probability of kill")+
  theme(legend.key=element_rect(fill="white"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))
ggsave(here("figures/canopy.conf.png"),dpi=300)
```

```{r graph effect size}
apatheme=theme_minimal()+
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        legend.title=element_blank(), 
        axis.text=element_text(size=10),
        axis.title=element_text(size=10),
        panel.background = element_rect(fill="white"))
plot_summs(elk,mule,model.names = c("Elk","Mule deer"),
           coefs=c("Elevation"="elevation",
                   "Canopy cover"="canopy_cover",
                   "Slope"="slope"))+
  apatheme+
  labs(x="Effect size",y=NULL)
ggsave(here("figures/effect.size.png"),dpi=300)
```

```{r FINAL prediction plots}
library(ggeffects)
elev<-ggpredict(elk1, terms="elevation [all]")
elev$species<-"Elk"
elev1<-ggpredict(mule1, terms="elevation [all]")
elev1$species<-"Mule deer"
elev.bind<-bind_rows(elev,elev1)
ggplot(elev.bind, aes(x, predicted)) + 
  geom_ribbon(data=elev,aes(ymin=conf.low, ymax=conf.high), alpha=0.15,fill="#436658")+
  geom_ribbon(data=elev1,aes(ymin=conf.low, ymax=conf.high), alpha=0.15,fill="#b4673b")+
  geom_smooth(aes(col=species))+
  scale_color_manual(values = c("#436658","#b4673b"), 
                     name = "",labels=c("Elk","Mule deer"))+
  labs(x="Elevation", y="Probability of kill")+
  theme(legend.key = element_rect(fill="white"),
        legend.position="top",
        legend.box.spacing = unit(0,"in"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))
ggsave(here("figures/elevation.bind1.png"),width=4,height=3,units="in",dpi=300)

canopy<-ggpredict(elk1, terms="canopy_cover [all]")
canopy$species<-"Elk"
canopy1<-ggpredict(mule1, terms="canopy_cover [all]")
canopy1$species<-"Mule deer"
canopy.bind<-bind_rows(canopy,canopy1)
ggplot(canopy.bind, aes(x, predicted)) + 
  geom_ribbon(data=canopy,aes(ymin=conf.low, ymax=conf.high), alpha=0.15,fill="#436658")+
  geom_ribbon(data=canopy1,aes(ymin=conf.low, ymax=conf.high), alpha=0.15,fill="#b4673b")+
  geom_smooth(aes(col=species))+
  scale_color_manual(values = c("#436658","#b4673b"), 
                     name = "",labels=c("Elk","Mule deer"))+
  labs(x="Canopy cover", y="Probability of kill")+
  theme(legend.key = element_rect(fill="white"),
        legend.position="top",
        legend.box.spacing = unit(0,"in"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))
ggsave(here("figures/canopy.bind1.png"),width=4,height=3,units="in",dpi=300)

slope<-ggpredict(elk1, terms="slope [all]")
slope$species<-"Elk"
slope1<-ggpredict(mule1, terms="slope [all]")
slope1$species<-"Mule deer"
slope.bind<-bind_rows(slope,slope1)
ggplot(slope.bind, aes(x, predicted)) + 
  geom_ribbon(data=slope,aes(ymin=conf.low, ymax=conf.high), alpha=0.15,fill="#436658")+
  geom_ribbon(data=slope1,aes(ymin=conf.low, ymax=conf.high), alpha=0.15,fill="#b4673b")+
  geom_line(aes(col=species),size=1)+
  scale_color_manual(values = c("#436658","#b4673b"), 
                     name = "",labels=c("Elk","Mule deer"))+
  labs(x="Slope", y="Probability of kill")+
  theme(legend.key = element_rect(fill="white"),
        legend.position="top",
        legend.box.spacing = unit(0,"in"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))
ggsave(here("figures/slope.bind1.png"),width=4,height=3,units="in",dpi=300)
```

```{r raw data plots}
dat.plot<-data.frame(elev.bin=seq(1500,2800,100))%>%
  mutate(row=as.numeric(row.names(.)))
preplot<-premod
preplot$row<-findInterval(premod$elevation,dat.plot$elev.bin)
preplot<-left_join(preplot,dat.plot,by="row")%>%
  dplyr::select(-row)

dat.plot1<-data.frame(slope.bin=seq(0,55,5))%>%
  mutate(row=as.numeric(row.names(.)))
preplot$row<-findInterval(preplot$slope,dat.plot1$slope.bin)
preplot<-left_join(preplot,dat.plot1,by="row")%>%
  dplyr::select(-row)

dat.plot2<-data.frame(canopy.bin=seq(0,70,5))%>%
  mutate(row=as.numeric(row.names(.)))
preplot$row<-findInterval(preplot$canopy_cover,dat.plot2$canopy.bin)
preplot<-left_join(preplot,dat.plot2,by="row")%>%
  dplyr::select(-row)

elevplot<-preplot%>%
  filter(kill==1)%>%
  group_by(elev.bin, species)%>%
  summarize(nobs=n())
slopeplot<-preplot%>%
  filter(kill==1)%>%
  group_by(slope.bin, species)%>%
  summarize(nobs=n())
canopyplot<-preplot%>%
  filter(kill==1)%>%
  group_by(canopy.bin, species)%>%
  summarize(nobs=n())

ggplot(elevplot)+
  geom_bar(aes(x=as.factor(elev.bin),y=nobs,fill = species),  
           stat = "identity", position = "dodge")+
  scale_fill_manual(values = c("#436658","#b4673b"), 
                     name = "",labels=c("Elk","Mule deer"))+
  labs(x="Elevation (m)", y="Number of kills")+
  theme(legend.position="top",
        legend.box.spacing = unit(0,"in"))
ggsave(here("figures/bar.elevation.png"), width=4, height=3, 
            units="in", dpi=300)

ggplot(slopeplot)+
  geom_bar(aes(x=as.factor(slope.bin),y=nobs,fill = species),  
           stat = "identity", position = "dodge")+
  scale_fill_manual(values = c("#436658","#b4673b"), 
                     name = "",labels=c("Elk","Mule deer"))+
  labs(x="Slope", y="Number of kills")+
  theme(legend.position="top",
        legend.box.spacing = unit(0,"in"))
ggsave(here("figures/bar.slope.png"), width=4, height=3, 
            units="in", dpi=300)

ggplot(canopyplot)+
  geom_bar(aes(x=as.factor(canopy.bin),y=nobs,fill = species),  
           stat = "identity", position = "dodge")+
  scale_fill_manual(values = c("#436658","#b4673b"), 
                     name = "",labels=c("Elk","Mule deer"))+
  labs(x="Canopy cover (%)", y="Number of kills")+
  theme(legend.position="top",
        legend.box.spacing = unit(0,"in"))
ggsave(here("figures/bar.canopy.png"), width=4, height=3, 
            units="in", dpi=300)
```