---
title: "prelim_model"
author: "Tam Ta"
date: "4/19/2022"
output: html_document
---

```{r load library, include=FALSE}
library(here)
library(tidyverse)
library(ggplot2)
library(lme4)
```

```{r load data, include=FALSE}
dir_all<-read.csv(here("data/all_elev.csv"))
  #subset(species=="elk"|species=="mule deer")
```

```{r plot}
ggplot(dir_all)+
         #filter(species=="elk"|species=="mule deer"))+
  geom_boxplot(aes(x=species,y=elevation))+
  theme(axis.text.x=element_text(angle=45, hjust=1))

ggplot(dir_mod)+
  geom_boxplot(aes(x=species,y=elevation))+
  geom_jitter(aes(x=species,y=elevation),shape=16,
              position=position_jitter(0.2))+
  xlab("Species")+
  ylab("Elevation (m)")+
  scale_x_discrete(labels=c("Elk","Mule deer"))
ggsave(here("figures/elev_species_box.png"),width=5,height=3,
       units="in",dpi=300)

ggplot(dir_all)+
  geom_boxplot(aes(x=season,y=elevation))
ggplot(dir_all%>%
         filter(species=="elk"|species=="mule deer"))+
  geom_point(aes(x=jday,y=elevation,color=species))
ggplot(dir_all%>%
         filter(species=="elk"|species=="mule deer")%>%
         filter(age!="unk"))+
  geom_boxplot(aes(x=age,y=elevation))
ggplot(dir_all%>%
         filter(species=="elk"|species=="mule deer")%>%
         filter(age!="unk"))+
  geom_point(aes(x=jday,y=elevation,color=age))

ggplot(dir_bin)+
  geom_bar(aes(x=elev.bin,y=nobs,fill=as.factor(ELK)),
           stat="identity",position=position_dodge(0.9))+
  theme(axis.text.x=element_text(angle=45, hjust=1))+
  xlab("Elevation (m)")+
  ylab("Number of observations")+
  scale_fill_manual(values = c("#39638b", "#26a185"), 
                    name = "Species", labels=c("Mule deer",
                                               "Elk"))
ggsave(here("figures/elev_species.png"),width=5,height=3,
       units="in",dpi=300)
```
Looks like it's not a linear relationship but exponential. Therefore, we will perform a generalized linear model.

```{r change data to binary}
dir_mod<-dir_all%>%
  subset(dir_all$season=="winter")
dir_mod<-dir_mod%>%
  subset(dir_mod$species=="mule deer" | 
           dir_mod$species=="elk")%>%
  mutate(MUDE=ifelse(species=="mule deer",1,0))%>%
  mutate(ELK=ifelse(species=="elk",1,0))%>%
  separate(elevation,
           into=c("elev.bin","ext"),
           sep=2, remove=FALSE)%>%
  select(-ext)%>%
  filter(!is.na(season))
dir_mod$elev.bin<-paste0(dir_mod$elev.bin, "00")
dir_bin<-dir_mod%>%
  group_by(elev.bin,ELK,MUDE)%>%
  summarize(nobs=n())
```
The models that we will run once we have complete dataset from GIS team:
name<-glm(ELK~elevation + distance to water + distance to roads + canopy cover)

name<-glm(MUDE~elevation + distance to water + distance to roads + canopy cover)

name<-glm(age.n~elevation + distance to water + distance to roads + canopy cover)

```{r model for elk}
m.elk<-glm(ELK~elevation,data=dir_mod,family=binomial)
summary(m.elk)
```
Elevation does not have a significant effect on the probability of getting an elk.
Compared to adult, old adult, yearling, and young elk has a significantly higher probability of being killed.

```{r model for mule deer}
m.mude<-glm(MUDE~elevation,data=dir_mod,family="binomial")
summary(m.mude)
```

```{r model for mule deer}
ele.prho<-glm(PRHO~elevation,data=dir_mod,family="binomial")
summary(ele.prho)
```

```{r model for mule deer}
ele.mamo<-glm(MAMO~elevation,data=dir_mod,family="binomial")
summary(ele.mamo)
```

```{r model for mule deer}
ele.bish<-glm(BISH~elevation,data=dir_mod,family="binomial")
summary(ele.bish)
```

